Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-10-21 pkg/disttree/inst/draft/draft.Rnw
further revisions based on reviews
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2348 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: f085d88cc0b20e11401fa7b30d38ad5601ed0a2f

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}
In the following, particular focus is given to distributional forests as a method for obtaining
probabilistic forecasts by leveraging the strengths of random forests: the ability to
capture both smooth and abruptly changing functions along with simultaneous selection of
variables and possibly complex interactions. Thus, these properties make the method particularly
appealing in case of many covariates with unknown effects and interactions where it would be 
challenging to specify a distributional regression model like GAMLSS.

@@

\subsubsection*{Variable Selection}
\hlc{Generally, there are many possibilities how to specify the variables that are
to be included in a model. Especially for a low number of variables all types
of interactions can be employed and mixed effect models and separations into 
main effects and interaction effects can be considered allowing for easier 
interpretations} (see for example \cite{Wood+Scheipl+Faraway:2013}, 
\cite{Goicoa+Adin+Ugarte:2018}, \cite{Ugarte+Adin+Goicoa:2017}).
\hlc{However, in the application presented in} Section~\ref{sec:precipitation}
\hlc{altogether 80 variables are available. Even if only paired interactions were
considered this would already lead to 3160 additional covariates in the model
making it very difficult to keep the model manageable regarding computational
costs but also interpretability. Therefore, only main effects are considered
for the distributional forest and the boosted GAMLSS which both automatically
select covariates while meteorological expert knowledge was applied to 
select covariates and specify interactions for the prespecified GAMLSS.}

@@

Alternatively, a distribution including an additional parameter capturing skewness
\citep{Scheuerer+Hamill:2015, Baran+Nemoda:2016} could be applied. For example, 
in transformation forests} \citep{Hothorn+Zeileis:2017} \hlc{a very flexible 
five-parametric distribution family is employed such that skewness can be modeled 
explicitly.
Evaluating further distributions might be of interest but this would go beyond the
scope of this manuscript.
}

>>>>

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}

\hlc{However, distributional forests should not be considered as a 
replacement of GAMLSS but rather as a complementing method in the field of
distributional models, similar to the way how GAM and random forests are complements 
in the conditional mean regression toolbox.}

In the following, particular focus is given to distributional forests as a method for
obtaining probabilistic forecasts by leveraging the strengths of random forests: the 
ability to capture both smooth and abruptly changing functions along with simultaneous selection of variables and possibly complex interactions. Thus, these properties make the method particularly appealing in case of many covariates with unknown effects and 
interactions where it would be challenging to specify a distributional regression model 
like GAMLSS.

@@

\subsubsection*{Variable Selection}
\hlc{Generally, there are many possibilities how to specify the variables that are
to be included in a model. Especially for a low number of variables all types
of interactions can be employed and mixed effect models and separations into 
main effects and interaction effects can be considered allowing for easier 
interpretations} (see for example \cite{Wood+Scheipl+Faraway:2013}, 
\cite{Goicoa+Adin+Ugarte:2018}, \cite{Ugarte+Adin+Goicoa:2017}).
\hlc{However, in the application presented in} Section~\ref{sec:precipitation}
\hlc{altogether 80 variables are available. Even if only paired interactions were
considered this would already lead to 3160 additional covariates in the model
making it very difficult to keep the model manageable regarding computational
costs but also interpretability. Therefore, only main effects are considered
for the distributional forest and the boosted GAMLSS which both automatically
select covariates while meteorological expert knowledge was applied to 
select covariates and specify interactions for the prespecified GAMLSS.

As mentioned before, distributional forests should not be seen as a replacement
for GAMLSS but rather as a complementary method. GAMLSS provide a powerful framework
in the field of distributional modeling as they allow for the combination of 
interpretability and a high goodnes-of-fit. For distributional forests, the focus
is mainly on forecasting skill, automatic variable selection and capturing also 
non-additive effects rather than on interpretability. Therefore, both approaches
provide valuable features and advantages that can complement each other in many 
applications as it can not be expected that one of them is in general better than 
the other. 
}

@@

Alternatively, a distribution including an additional parameter capturing skewness}
\citep{Scheuerer+Hamill:2015, Baran+Nemoda:2016} \hlc{could be applied. For example, 
in transformation forests} \citep{Hothorn+Zeileis:2017} \hlc{a very flexible 
five-parametric distribution family is employed such that skewness can be modeled 
explicitly.
Evaluating further distributions might be of interest but this would go beyond the
scope of this manuscript.
}
